---
title: "Clinical Trials"
author: "Rafael Sommer"
date: "05/05/2020"
output: html_document
---
# Clinical Trials common used scripts
This document will contain code that exemplifies common clinical trial practices as well as templates for personal designs. I used some of the Clinical Trial DataCamp course to extract some of the code, and libraries.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo = FALSE}
library(compareGroups)
library(ggplot2)
library(blockrand)
```
Libraries used and its version
- Package compareGroups version 4.4.1
- Package blockrand version 1.5

# Randomization
## Simple and block randomization
This do not take any extra factors in account to assign groups
```{r}
treatment <- c("A","B")
# list <- sample(treatment, x, replace = TRUE)
# cat(list,sep="\n") # Command to show concatenated list
```
This method, however, can generate unbalanced group for small sample size, in order to correct for that, we may choose a block randomization design:
```{r}
# Template
num = 10 # Minimum number of subjects
numArm = 2 # Number of treatments arms
x = 2 # Block size (randomizes x * numArm)
y = 2 # Block size (same, you can use several sizes for random diferent blocks)
block.list <- blockrand(n = num, num.levels = numArm, block.sizes = c(x,y))

# E.G
block.list <- blockrand(n = 50, num.levels = 2, block.sizes = c(2,2))
head(block.list)
# To make the end of each block unpredictable, you may allow blocks to have different sizes at random, for eg:
block.list <- blockrand(n = 20, num.levels = 2, block.sizes = c(1,2))
# You can see that some blocks have a size = 2 (n*1) or 4 (n*2)
block.list
```
## Stratified randomization
This is helpful if you want to balance the treatment arms for certain known variable, randomization carries a different block for each stratified group,

 


# Sample size and power calculation






# Dataset Exploration
We will use the compareGroups library which facilitate comparing for baseline group characteristics and building tables. Let´s do a basic baseline comparation example:
```{r, echo = TRUE}
# Template
#baseline_info <- compareGroups(group ~ x1 + x2 + ... + xn, data = dataset)
#baseline_table <- createTable(baseline_info, show.ratio = F, show.p.overall = F)
#baseline_table
```