# Probability with R

Author: Rafael Sommer

### This script provide information and R scripts that I used to study probability. Also, some of the scripts may be helpful when dealing with bayesian statistics and generating data

### Binomial distribution
```{r}
set.seed(1234)
# Coin flip
rbinom (n = 1, size = 1, prob = .5) 
# rbinom generates binomial distributions with:
# n = number of coins flipped, so:
rbinom (n = 10, size = 1, prob = .5)

# size = number of coins that should be flipped within each draw, so:
rbinom (n = 1, size = 4, prob = .5)

# prob = probability of getting a 1 (heads), unfair coin
rbinom (n = 10, size = 1, prob = .9)

```
The binomial distribution is represented by the following notation:

$X \sim {\sf Binom}(n, \pi)$ , 

Where $n$ = size and $\pi$ = probability

### Density and cumulative density

dbinom function estimate the probability density for x "sucessess"
```{r} 
dbinom (5, 10, 0.5) 
```

This meaning that is a coin with p = .5 is flipped 10 times, 0.24 of those times there will be 5 (x) sucessess. We can confirm it by simulating 10.000 trials of 10 flips and extracting the mean of the sucessess (given that sucess is 1)
```{r}
head(rbinom(10000,10,.5))
mean (rbinom(10000,10,0.5) == 5)
```

For cumulative densities we are interested in the probability of at least "q" sucesses
```{r}
pbinom(5, 10, 0.5, lower.tail = TRUE)
```
This gives us the probability that are AT LEAST 5 sucessess from 10 trials with sucess probability of 0.5. We can also figure the probability of having MORE than 5 sucesses from 10 trials, which is complementary 
```{r}
pbinom(5, 10, 0.5, lower.tail = FALSE)
```

